---
title: "Multilevel Models"
author: 
- "Mine Dogucu"
- "Noah Johnson"
date: "April 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(install.load)
install_load('tidyverse')
install_load('lme4')
```

We will continue to use the hsb.csv file. In the last class, we had group mean centered the ses variable.

```{r}
hsb <- read.csv("hsb.csv")

group_means <- hsb %>% 
  group_by(schoolid) %>% 
  summarize(group.mean.ses=mean(ses))

hsb <- merge(hsb, group_means, by="schoolid")

hsb$group.mean.cen.ses <- hsb$ses - hsb$group.mean.ses
```

## Random Coefficients Model

```{r}
model2 <- lmer(mathach ~ group.mean.cen.ses + (group.mean.cen.ses | schoolid), data=hsb)
model2.sum <- summary(model2)
model2.sum
```

### Model Notation

$mathach_{ij} = \beta_{0j} + \beta_{1j}*(ses_{ij} - \bar{ses}_{.j}) + r_{ij}$

$\beta_{0j} = \gamma_{00} + u_{0j}$

$\beta_{1j} = \gamma_{10} + u_{1j}$

### Parameters

$\gamma_{00} = `r model2.sum$coefficients[1,1]`$ is the estimated grand mean math score.

$\gamma_{10} = `r model2.sum$coefficients[2,1]`$ is the average effect of hedres on math score.

$var(r_{ij}) = 36.7$ is the within-group variance.

$var(u_{0j}) = 8.681$ is the between-group variance of the intercept $\beta_{0j}$.

$var(u_{1j}) = 0.694$ is the between-group variance of the slope $\beta_{1j}$.

## Contextual Model

Sector = 0 for public schools, 1 for catholic schools

```{r}
model3 <- lmer(mathach ~ group.mean.cen.ses * sector + (group.mean.cen.ses | schoolid), data=hsb)
model3.sum <- summary(model3)
model3.sum
```

### Model Notation
Level 1: $mathach_{ij} = \beta_{0j} + \beta_{1j}*(ses_{ij} - \bar{ses_{.j}}) + r_{ij}$

Level 2: 

$\beta_{0j} = \gamma_{00} + \gamma_{01}*sector_j + u_{0j}$

$\beta_{1j} = \gamma_{10} + \gamma_{11}*sector_j + u_{1j}$

Mixed: 

$$
\begin{aligned}
mathach_{ij} = \gamma_{00} + \gamma_{01}*sector_j + u_{0j} + \gamma_{10}(ses_{ij} - \bar{ses_{.j}}) + 
\\
\gamma_{11} * sector_j * (ses_{ij} - \bar{ses_{.j}}) + u_{1j} (ses_{ij} - \bar{ses_{.j}}) + r_{ij}
\end{aligned}
$$

### Parameters

$\gamma_{00} = `r model3.sum$coefficients[[1,1]]`$. Expected math achievement score for public schools.

$\gamma_{01} = `r model3.sum$coefficients[[2,1]]`$. Expected increase in expected math achievement for catholic schools.

$\gamma_{10} = `r model3.sum$coefficients[[3,1]]`$. The effect of ses on math achievement for public schools.

$\gamma_{11} = `r model3.sum$coefficients[[4,1]]`$. The increase of the effect of SES on math achievement for catholic schools.

$s^2 = `r model3.sum$sigma^2`$. The residual within-group variance

$\tau_{00} = 6.7378$. The between-group variance in the expected math achievement.

$\tau_{11} = 0.2657$. The between-group variance in the effect of ses.