---
title: "Multilevel Models"
author: "Mine Dogucu"
date: "April 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(install.load)
install_load('tidyverse')
install_load('lme4')
```

We will continue to use the hsb.csv file. In the last class, we had group mean centered the ses variable.

```{r}
hsb <- read.csv("hsb.csv")
hsb.sum <- hsb %>% 
  group_by(schoolid) %>% 
  summarize(group.mean.ses=mean(ses))
hsb <- merge(hsb, hsb.sum, by="schoolid")
hsb$group.mean.cen.ses <- hsb$ses - hsb$group.mean.ses
```

## Random Coefficients Model

```{r}
model2 <- lmer(mathach ~ group.mean.cen.ses + (group.mean.cen.ses | schoolid), data=hsb)
summary(model2)
```

### Model Notation

### Parameters

## Contextual Model

Sector = 0 for public schools,
1 for catholic schools

```{r}
model3 <- lmer(mathach ~ group.mean.cen.ses * sector + (group.mean.cen.ses | schoolid), data=hsb)
model3.sum <- summary(model3)
model3.sum
```

### Model Notation
Level 1: $Y_{ij} = \beta_{0j} + \beta_{1j}(ses_{ij} - \bar{ses_{.j}}) + r_{ij}$

Level 2: $\beta_{0j} = \gamma_{00} + \gamma_{01}*sector_j + u_{0j}$

$\beta_{0j} = \gamma_{00} + \gamma_{01}*sector_j + u_{0j}$

Mixed: $Y_{ij} = \gamma_{00} + \gamma_{01}*sector_j + u_{0j} + \gamma_{10}(ses_{ij} - \bar{ses_{.j}}) + \gamma_{11} * sector_j (ses_{ij} - \bar{ses_{.j}}) + u_{1j} (ses_{ij} - \bar{ses_{.j}}) + r_{ij}$

### Parameters

$\gamma_{00} = `r model3.sum$coefficients[[1,1]]`$. Expected math achievement score for public schools.

$\gamma_{01} = `r model3.sum$coefficients[[2,1]]`$. Expected increase in expected math achievement for catholic schools.

$\gamma_{10} = `r model3.sum$coefficients[[3,1]]`$. The effect of ses on math achievement for public schools.

$\gamma_{11} = `r model3.sum$coefficients[[4,1]]`$. The increase of the effect of SES on math achievement for catholic schools.

$\sigma^2 = `r model3.sum$sigma^2`$ Within-group variance

$\tau_{00} = `r `$

$\tau_{11} = `r `$
of between school variability in the effect of ses can be explained by sector